Meta analysis of differential expression results from AMP-AD (Control vs AD)

### 2016-01-27 code/Rmd/Mayo_Cerebellum_RNASeq.Rmd
Minor modifications

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Fix metadata
METADATA$AgeAtDeath = gsub("_or_above", "", METADATA$AgeAtDeath)
METADATA <- METADATA %>%
  dplyr::filter(!(SampleID %in% c("742_CER", "11386_CER","11311_CER","1923_CER","11477_CER")))
```
Following five samples were considered outliers and removed 742\_CER, 11386\_CER, 11311\_CER, 1923\_CER, 11477\_CER
```{r preprocess.data1}
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum RNASeq', parentId = parentId)
CODE <- synStore(CODE)

>>>>

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Fix metadata
METADATA$AgeAtDeath = gsub("_or_above", "", METADATA$AgeAtDeath)
METADATA$Diagnosis = gsub("Pathologic Aging", "Pathologic_Aging", METADATA$Diagnosis)
METADATA <- METADATA %>%
  dplyr::filter(!(SampleID %in% c("742_CER", "11386_CER","11311_CER","1923_CER","11477_CER")))
```
Following five samples were considered outliers and removed 742\_CER, 11386\_CER, 11311\_CER, 1923\_CER, 11477\_CER
```{r preprocess.data1}
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp}
# Fit contrast
contrast = makeContrasts(contrasts=c("DiagnosisControl-DiagnosisAD",
                                     "DiagnosisControl-DiagnosisPathologic_Aging",
                                     "DiagnosisControl-DiagnosisPSP",
                                     "DiagnosisPathologic_Aging-DiagnosisAD",
                                     "DiagnosisPathologic_Aging-DiagnosisAD",
                                     "DiagnosisPSP-DiagnosisAD"),
                         levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT.CONTR = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = topTable(FIT.CONTR, coef=1, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1])
DE = rownameToFirstColumn(DE, 'ensembl_gene_id') %>%
  left_join(Ensemble2HGNC)
p <- ggplot(DE, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05))
p

tmp = DE %>% dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(hgnc_symbol) %>%
  unlist %>% unique
kable(data.frame(Comparison = "Control_vs_AD",
           nDiffExp = length(tmp),
           DiffExpGenes = paste(tmp, collapse = ',')))  
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum RNASeq', parentId = parentId)
CODE <- synStore(CODE)

### 2016-01-27 code/Rmd/MetaAnalysis.Rmd
Minor modifications

# Load required libraries
library(synapseClient)
library(RColorBrewer)
library(ggplot2)
library(data.table)
library(plyr)
library(dplyr)
library(psych)
library(tidyr)

@@

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names = T)
tmp = sapply(file.sources,source,.GlobalEnv)

@@

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 7/8 datasets')
rownames(Adj.P.Value)[which(rowSums(Adj.P.Value<=0.05)==7)]

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 6/8 datasets')
rownames(Adj.P.Value)[which(rowSums(Adj.P.Value<=0.05)==6)]
```

@@

# Store pvalues
tmp = join_all(list(Pmin,Pfb))
write.table(tmp, file = './Pval.tsv', sep = '\t', row.names=F, quote=F)
FILE_OBJ = File("./Pval.tsv", name = 'Meta Differential Expression Analysis FDR', parentId = CODE$properties$id)
FILE_OBJ = synStore(FILE_OBJ, 
                    used = as.character(c(EXP_ID, COV_ID)),
                    executed = thisFile,
                    activityName = 'Differential expression analysis of control-AD')

# Store differentially expressed genes
write.table(rownameToFirstColumn(Adj.P.Value, 'hgnc_symbol'), file = './DiffGenes.tsv', sep = '\n', row.names=F, quote=F, col.names = F)
FILE_OBJ = File("./DiffGenes.tsv", name = 'Differentially expressed genes FDR (individual studies)', parentId = CODE$properties$id)
FILE_OBJ = synStore(FILE_OBJ, 
                    used = as.character(c(EXP_ID, COV_ID)),
                    executed = thisFile,
                    activityName = 'Differential expression analysis of control-AD')
```

>>>>

# Load required libraries
library(synapseClient)
library(RColorBrewer)
library(ggplot2)
library(data.table)
library(plyr)
library(dplyr)
library(psych)
library(tidyr)
library(Matrix)

@@

# source utility files from ../R/lib folder
source('../R/lib/rownameToFirstColumn.R')
# file.sources = list.files('../R/lib/rownfile.sources',pattern="row*.R", full.names = T)
# tmp = sapply(file.sources,source,.GlobalEnv)

@@

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 7/8 datasets')
paste(rownames(Adj.P.Value)[which(rowSums(Adj.P.Value<=0.05)==7)], collapse = ',')

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 6/8 datasets')
paste(rownames(Adj.P.Value)[which(rowSums(Adj.P.Value<=0.05)==6)], collapse = ',')
```

@@

# Store pvalues
tmp = join_all(list(Pmin,Pfb))
write.table(tmp, file = './Pval.tsv', sep = '\t', row.names=F, quote=F)
FILE_OBJ = File("./Pval.tsv", name = 'Meta Differential Expression Analysis FDR', parentId = CODE$properties$id)
FILE_OBJ = synStore(FILE_OBJ, 
                    used = as.character(DEXP_ID),
                    executed = thisFile,
                    activityName = 'Differential expression analysis of control-AD')

# Store differentially expressed genes
write.table(rownameToFirstColumn(Adj.P.Value, 'hgnc_symbol'), file = './DiffGenes.tsv', sep = '\n', row.names=F, quote=F, col.names = F)
FILE_OBJ = File("./DiffGenes.tsv", name = 'Differentially expressed genes FDR (individual studies)', parentId = CODE$properties$id)
FILE_OBJ = synStore(FILE_OBJ, 
                    used = as.character(DEXP_ID),
                    executed = thisFile,
                    activityName = 'Differential expression analysis of control-AD')
```
